---
title: "Loop extrusion simulations"
output: html_notebook
---

```{r}
library(foreach)
library(doParallel)
registerDoParallel(cores=8)
library(ggplot2)
library(dplyr)
library(IRanges)
library(RColorBrewer)
source("extrusion-sim-fn.R")
```


# Set up a simple demonstration simulation with:
- A 20kb fiber 
- 2 cohesins with processivity=10kb
- 3 CTCFs with a 90% capture rate
```{r}
set.seed(140)
sim <- Sim(2, polymer_length_bp = 20e3, processivity_bp = 10e3, bead_size = 1000, ctcf_capture_rate=0.9, ctcf_stabilization_factor = 2)
sim <- set_ctcfs(sim, orientation = "+", pos_bp=1500, occupancy_rate = 1)
sim <- set_ctcfs(sim, orientation = "+", pos_bp=1505, occupancy_rate = 1)
sim <- set_ctcfs(sim, orientation = "-", pos_bp=12500, occupancy_rate = 1)
sim
```

Visualize the initial state
```{r}
plot_fiber(sim)
```

# Advance 30 steps
```{r}
for (i in 1:30) {
    sim <- advance(sim)
    plot_fiber(sim)
}
```

# Simulate two 500kb-separated points with two surrounding CTCFs

```{r}
ctcf_occ_rate <- 0.5
sim <- Sim(8, polymer_length_bp = 240e3*8, processivity = 150e3, bead_size = 1000)
sim <- sim %>% set_ctcfs(orientation = "+", pos_bp=695e3, occupancy_rate = ctcf_occ_rate) 
sim <- sim %>% set_ctcfs(orientation = "-", pos_bp=1205e3, occupancy_rate = ctcf_occ_rate) 

pointA <- 700
pointB <- 1200

sim_df <- foreach(run = 1:3, .combine=rbind) %dopar% {
    foreach(i = 1:1000, .combine=rbind) %do% {
        sim <- advance(sim)
        data.frame(run=run, i=i, effective_dist = effective_dist(sim, pointA, pointB))
    }
}

sim_df %>% ggplot(aes(i, effective_dist, group=run)) + geom_line() + theme_bw()

```





# Simulate two 500kb-separated points with and without surrounding CTCFs

```{r}
iter <- 50
reps <- 5
    
system.time({
    ctcf_occ_rate <- 0.5
    sim <- Sim(8, polymer_length_bp = 240e3*8, processivity = 150e3, bead_size = 1000)
    
    sim_with_1_left_ctcf <- sim %>% set_ctcfs(orientation = "+", pos_bp=695e3, occupancy_rate = ctcf_occ_rate) 
    
    sim_with_4_left_ctcf <- sim %>% set_ctcfs(orientation = "+", pos_bp=692e3, occupancy_rate = ctcf_occ_rate) %>%
        set_ctcfs(orientation = "+", pos_bp=693e3, occupancy_rate = ctcf_occ_rate) %>%
        set_ctcfs(orientation = "+", pos_bp=694e3, occupancy_rate = ctcf_occ_rate) %>%
        set_ctcfs(orientation = "+", pos_bp=695e3, occupancy_rate = ctcf_occ_rate)
    
    sim_with_4_left_and_4_right_ctcf <- sim %>% set_ctcfs(orientation = "+", pos_bp=692e3, occupancy_rate = ctcf_occ_rate) %>%
        set_ctcfs(orientation = "+", pos_bp=693e3, occupancy_rate = ctcf_occ_rate) %>%
        set_ctcfs(orientation = "+", pos_bp=694e3, occupancy_rate = ctcf_occ_rate) %>%
        set_ctcfs(orientation = "+", pos_bp=695e3, occupancy_rate = ctcf_occ_rate) %>%
        set_ctcfs(orientation = "-", pos_bp=1205e3, occupancy_rate = ctcf_occ_rate) %>%
        set_ctcfs(orientation = "-", pos_bp=1206e3, occupancy_rate = ctcf_occ_rate) %>%
        set_ctcfs(orientation = "-", pos_bp=1207e3, occupancy_rate = ctcf_occ_rate) %>%
        set_ctcfs(orientation = "-", pos_bp=1208e3, occupancy_rate = ctcf_occ_rate)
    
    
    dist <- rbind((avg_effective_dist(sim, type="pairwise", pointA=700, pointB=1200, num_iter=iter, num_reps=reps) %>% cbind(CTCFs="_None")),
                  (avg_effective_dist(sim_with_1_left_ctcf, type="pairwise", pointA=700, pointB=1200, num_iter=iter, num_reps=reps) %>% cbind(CTCFs="1_left")),
                  (avg_effective_dist(sim_with_4_left_ctcf, type="pairwise", pointA=700, pointB=1200, num_iter=iter, num_reps=reps) %>% cbind(CTCFs="4_left")),
                  (avg_effective_dist(sim_with_4_left_and_4_right_ctcf, type="pairwise", pointA=700, pointB=1200, num_iter=iter, num_reps=reps) %>% cbind(CTCFs="4_left_4_right")))

})

```

```{r}
dist
```


```{r}
dist %>% ggplot(aes(CTCFs, effective_dist)) + geom_boxplot(color="grey75") + geom_jitter(width = 0.25) + theme_bw() 
```


The code below explores the required burn-in time and stability.

# Check simulation stability - No CTCFs
```{r}
sim <- Sim(8, polymer_length_bp = 240e3*8, processivity = 150e3, bead_size = 1000)

pointA <- 700
pointB <- 1200

sim_df <- foreach(run = 1:8, .combine=rbind) %dopar% {
    foreach(i = 1:50000, .combine=rbind) %do% {
        sim <- advance(sim)
        data.frame(run=run, i=i, effective_dist = effective_dist(sim, pointA, pointB))
    }
}
sim_df <- sim_df %>% group_by(run) %>% mutate(avg_effective_dist = cumsum(effective_dist)/i)
sim_df %>% ggplot(aes(i, avg_effective_dist, group=run)) + geom_line() + theme_bw()
#sim_df %>% filter(i>5000) %>% ggplot(aes(i, avg_effective_dist, group=run)) + geom_line() + theme_bw()
#saveRDS(sim_df, file="../data/sim_df.rds")
sim_df <- readRDS("../data/sim_df.rds")

```


```{r}

sim_df %>% filter(i>=150) %>% mutate(i=i-150) %>% group_by(run) %>% mutate(avg_effective_dist = cumsum(effective_dist)/i) %>% ggplot(aes(i, avg_effective_dist, group=run)) + geom_line() + theme_bw()

```


# Check simulation stability - With CTCFs
```{r}
sim <- sim_with_4_left_and_4_right_ctcf

pointA <- 700
pointB <- 1200

sim_df <- foreach(run = 1:4, .combine=rbind) %dopar% {
    foreach(i = 1:50000, .combine=rbind) %do% {
        sim <- advance(sim)
        data.frame(run=run, i=i, effective_dist = effective_dist(sim, pointA, pointB), 
                   num_extruded_beads=sum(sim@extruded_beads),
                    bound_plus_cohesins=sum(sim@plus_ctcfs$bound_cohesin),
                    bound_minus_cohesins=sum(sim@minus_ctcfs$bound_cohesin))
    }
}
saveRDS(sim_df, file="../data/sim_df_with_ctcf.rds")
sim_df <- readRDS("../data/sim_df_with_ctcf.rds")

```

# Effective length
```{r}
sim_df %>% filter(i>=150) %>% mutate(i=i-150) %>% group_by(run) %>% mutate(avg_effective_dist = cumsum(effective_dist)/i) %>% ggplot(aes(i, avg_effective_dist, group=run)) + geom_line() + theme_bw()

```
```{r}
sim_df %>% mutate(i_bin=5000*ceiling(i/5000)) %>% group_by(run, i_bin) %>% summarize(avg_effective_dist = mean(effective_dist))  %>% ggplot(aes(i_bin, avg_effective_dist, group=run)) + geom_line() + theme_bw() 

```


Why does it trend up?

# Num extruded beads
```{r}
sim_df %>% filter(i>=150) %>% mutate(i=i-150) %>% group_by(run) %>% mutate(avg_extruded_beads = cumsum(num_extruded_beads)/i) %>% ggplot(aes(i, avg_extruded_beads, group=run)) + geom_line() + theme_bw()

```




```{r}
sim_df %>% mutate(i_bin=5000*ceiling(i/5000)) %>% group_by(run, i_bin) %>% summarize(avg_extruded_beads = mean(num_extruded_beads))  %>% ggplot(aes(i_bin, avg_extruded_beads, group=run)) + geom_line() + theme_bw() 

```



# Num cohesin-bound CTCFs
```{r}
sim_df %>% mutate(i_bin=5000*ceiling(i/5000)) %>% group_by(run, i_bin) %>% summarize(bound_plus_cohesins = mean(bound_plus_cohesins),
                                                                                     bound_minus_cohesins = bound_minus_cohesins)  %>% 
    ggplot(aes(i_bin, bound_plus_cohesins, group=run)) + geom_line() + theme_bw() 

```
