---
title: "GIST FGF3 loop extrusion simulations"
output: html_notebook
---

```{r}
library(foreach)
library(doParallel)
registerDoParallel(cores=48)
library(ggplot2)
library(dplyr)
library(IRanges)
library(RColorBrewer)
library(stringr)
source("extrusion-sim-fn.R")
```


# Simulate the FGF3 locus

We're temporarily using 69,000kb-71,920kb as the 0 - 1,920 range
FGF3 TSS: 69,819,416kb (819.416kb)
Main K27ac peak: 69,922,000-69,928,000 (925kb)

name: MA0139.1
location: chr11:69643502-69643520 (-)
score: 13.082

name: MA0139.1
location: chr11:69670403-69670421 (-)
score: 20.9508

name: MA0139.1
location: chr11:69686029-69686047 (+)
score: 14.5738

name: MA0139.1
location: chr11:69704750-69704768 (+)
score: 17.459

name: MA0139.1
location: chr11:69718314-69718332 (+)
score: 12.4426

-------------

name: MA0139.1
location: chr11:69801712-69801730 (-)
score: 20.4918

name: MA0139.1
location: chr11:69842277-69842295 (-)
score: 20.0328

name: MA0139.1
location: chr11:69888726-69888744 (-)
score: 20.1475

name: MA0139.1
location: chr11:69910245-69910263 (+)
score: 21.2787

name: MA0139.1
location: chr11:69915619-69915637 (+)
score: 20.5082

name: MA0139.1
location: chr11:69919002-69919020 (+)
score: 21.0164

name: MA0139.1
location: chr11:69950264-69950282 (+)
score: 13.0


## Simulate without EP extrusion blocking

```{r}
rds <- "rds/fgf3-nt-b-c-coh8-occrate1-capturerate0.25-50kiter-100reps.rds"

ctcf_occ_rate <- 1
sim <- Sim(8, polymer_length_bp = 240e3*8, processivity_bp = 150e3, bead_size = 1000, ctcf_capture_rate=0.25)
sim <- sim %>% add_ctcf(orientation = "-", pos_bp=c(643502, 670403, 801712, 842277, 888726), occupancy_rate = ctcf_occ_rate) 
sim <- sim %>% add_ctcf(orientation = "+", pos_bp=c(686029, 704750, 718314, 910245, 915619, 919002, 950264), occupancy_rate = ctcf_occ_rate) 

pointA <- 819 # 819416
pointB <- 925

# Full distance
pointB - pointA

sim_nt <- sim
plot_fiber(sim_nt, 620, 950, main="NT")

sim_b <- sim_nt %>% delete_ctcf(pos_bp = c(910245, 915619, 919002))
plot_fiber(sim_b, 620, 950, main="B")

sim_c <- sim_nt %>% delete_ctcf(pos_bp = c(888726, 910245, 915619, 919002))
plot_fiber(sim_c, 620, 950, main="C")

#sim_noctcf <- sim_nt %>% delete_ctcf(pos_bp = c(801712, 842277, 888726, 910245, 915619, 919002, 950264))
#plot_fiber(sim_noctcf, 620, 950, main="C")

pointA <- 819 # 819416 FGF3P
#pointA <- 887 # Left of boundary
pointB <- 925

if (file.exists(rds)){
  df <- readRDS(rds)
} else {
  num_iter <- 50000
  burn_in <- 10000
  num_reps <- 100
  
  system.time({
    df <- rbind(
      cbind(effective_dist_sim(sim, type="pairwise", pointA=pointA, pointB=pointB, num_iter=num_iter, num_reps=num_reps, burn_in=burn_in), cond="NT"),
      cbind(effective_dist_sim(sim_b, type="pairwise", pointA=pointA, pointB=pointB, num_iter=num_iter, num_reps=num_reps, burn_in=burn_in), cond="B"),
      cbind(effective_dist_sim(sim_c, type="pairwise", pointA=pointA, pointB=pointB, num_iter=num_iter, num_reps=num_reps, burn_in=burn_in), cond="C"))
  })
  
  df$cond <- factor(df$cond, levels=c("NT", "B", "C"))
  saveRDS(df, file=rds)  
}

############################

rds <- "rds/fgf3-nt-b-c-coh12-occrate1-capturerate0.25-50kiter-100reps.rds"

ctcf_occ_rate <- 1
sim <- Sim(12, polymer_length_bp = 240e3*8, processivity_bp = 150e3, bead_size = 1000, ctcf_capture_rate=0.25)
sim <- sim %>% add_ctcf(orientation = "-", pos_bp=c(643502, 670403, 801712, 842277, 888726), occupancy_rate = ctcf_occ_rate) 
sim <- sim %>% add_ctcf(orientation = "+", pos_bp=c(686029, 704750, 718314, 910245, 915619, 919002, 950264), occupancy_rate = ctcf_occ_rate) 
pointA <- 819 # 819416
pointB <- 925
# Full distance
pointB - pointA
sim_nt <- sim
sim_b <- sim_nt %>% delete_ctcf(pos_bp = c(910245, 915619, 919002))
sim_c <- sim_nt %>% delete_ctcf(pos_bp = c(888726, 910245, 915619, 919002))
pointA <- 819 # 819416 FGF3P
#pointA <- 887 # Left of boundary
pointB <- 925
if (file.exists(rds)){
  df <- readRDS(rds)
} else {
  num_iter <- 50000
  burn_in <- 10000
  num_reps <- 100
  
  system.time({
    df <- rbind(
      cbind(effective_dist_sim(sim, type="pairwise", pointA=pointA, pointB=pointB, num_iter=num_iter, num_reps=num_reps, burn_in=burn_in), cond="NT"),
      cbind(effective_dist_sim(sim_b, type="pairwise", pointA=pointA, pointB=pointB, num_iter=num_iter, num_reps=num_reps, burn_in=burn_in), cond="B"),
      cbind(effective_dist_sim(sim_c, type="pairwise", pointA=pointA, pointB=pointB, num_iter=num_iter, num_reps=num_reps, burn_in=burn_in), cond="C"))
  })
  
  df$cond <- factor(df$cond, levels=c("NT", "B", "C"))
  saveRDS(df, file=rds)  
}

###

rds <- "rds/fgf3-nt-b-c-coh16-occrate1-capturerate0.25-50kiter-100reps.rds"

ctcf_occ_rate <- 1
sim <- Sim(16, polymer_length_bp = 240e3*8, processivity_bp = 150e3, bead_size = 1000, ctcf_capture_rate=0.25)
sim <- sim %>% add_ctcf(orientation = "-", pos_bp=c(643502, 670403, 801712, 842277, 888726), occupancy_rate = ctcf_occ_rate) 
sim <- sim %>% add_ctcf(orientation = "+", pos_bp=c(686029, 704750, 718314, 910245, 915619, 919002, 950264), occupancy_rate = ctcf_occ_rate) 
pointA <- 819 # 819416
pointB <- 925
# Full distance
pointB - pointA
sim_nt <- sim
sim_b <- sim_nt %>% delete_ctcf(pos_bp = c(910245, 915619, 919002))
sim_c <- sim_nt %>% delete_ctcf(pos_bp = c(888726, 910245, 915619, 919002))
pointA <- 819 # 819416 FGF3P
#pointA <- 887 # Left of boundary
pointB <- 925
if (file.exists(rds)){
  df <- readRDS(rds)
} else {
  num_iter <- 50000
  burn_in <- 10000
  num_reps <- 100
  
  system.time({
    df <- rbind(
      cbind(effective_dist_sim(sim, type="pairwise", pointA=pointA, pointB=pointB, num_iter=num_iter, num_reps=num_reps, burn_in=burn_in), cond="NT"),
      cbind(effective_dist_sim(sim_b, type="pairwise", pointA=pointA, pointB=pointB, num_iter=num_iter, num_reps=num_reps, burn_in=burn_in), cond="B"),
      cbind(effective_dist_sim(sim_c, type="pairwise", pointA=pointA, pointB=pointB, num_iter=num_iter, num_reps=num_reps, burn_in=burn_in), cond="C"))
  })
  
  df$cond <- factor(df$cond, levels=c("NT", "B", "C"))
  saveRDS(df, file=rds)  
}
```

## Simulate with EP extrusion blocking

```{r}
rds <- "rds/fgf3-nt-b-c-coh12-occrate1-capturerate0.5-epcapturerate0.25-50kiter-100reps.rds"

ctcf_occ_rate <- 1
sim <- Sim(12, polymer_length_bp = 240e3*8, processivity_bp = 150e3, bead_size = 1000, ctcf_capture_rate=0.5)
sim <- sim %>% add_ctcf(orientation = "-", pos_bp=c(643502, 670403, 801712, 842277, 888726), occupancy_rate = ctcf_occ_rate) 
sim <- sim %>% add_ctcf(orientation = "+", pos_bp=c(686029, 704750, 718314, 910245, 915619, 919002, 950264), occupancy_rate = ctcf_occ_rate) 

## Add E, P blocking
sim <- sim %>% add_ctcf(orientation = "+", pos_bp=c(819416), occupancy_rate = ctcf_occ_rate/2) 
sim <- sim %>% add_ctcf(orientation = "-", pos_bp=c(925000), occupancy_rate = ctcf_occ_rate/2) 

pointA <- 819 # 819416
pointB <- 925

# Full distance
pointB - pointA

sim_nt <- sim
plot_fiber(sim_nt, 620, 950, main="NT")

sim_b <- sim_nt %>% delete_ctcf(pos_bp = c(910245, 915619, 919002))
plot_fiber(sim_b, 620, 950, main="B")

sim_c <- sim_nt %>% delete_ctcf(pos_bp = c(888726, 910245, 915619, 919002))
plot_fiber(sim_c, 620, 950, main="C")

#sim_noctcf <- sim_nt %>% delete_ctcf(pos_bp = c(801712, 842277, 888726, 910245, 915619, 919002, 950264))
#plot_fiber(sim_noctcf, 620, 950, main="C")

pointA <- 819 # 819416 FGF3P
#pointA <- 887 # Left of boundary
pointB <- 925

if (file.exists(rds)){
  df <- readRDS(rds)
} else {
  num_iter <- 50000
  burn_in <- 10000
  num_reps <- 100
  
  system.time({
    df <- rbind(
      cbind(effective_dist_sim(sim, type="pairwise", pointA=pointA, pointB=pointB, num_iter=num_iter, num_reps=num_reps, burn_in=burn_in), cond="NT"),
      cbind(effective_dist_sim(sim_b, type="pairwise", pointA=pointA, pointB=pointB, num_iter=num_iter, num_reps=num_reps, burn_in=burn_in), cond="B"),
      cbind(effective_dist_sim(sim_c, type="pairwise", pointA=pointA, pointB=pointB, num_iter=num_iter, num_reps=num_reps, burn_in=burn_in), cond="C"))
  })
  
  df$cond <- factor(df$cond, levels=c("NT", "B", "C"))
  saveRDS(df, file=rds)  
}

############################

rds <- "rds/fgf3-nt-b-c-coh16-occrate1-capturerate0.5-epcapturerate0.25-50kiter-100reps.rds"

ctcf_occ_rate <- 1
sim <- Sim(16, polymer_length_bp = 240e3*8, processivity_bp = 150e3, bead_size = 1000, ctcf_capture_rate=0.5)
sim <- sim %>% add_ctcf(orientation = "-", pos_bp=c(643502, 670403, 801712, 842277, 888726), occupancy_rate = ctcf_occ_rate) 
sim <- sim %>% add_ctcf(orientation = "+", pos_bp=c(686029, 704750, 718314, 910245, 915619, 919002, 950264), occupancy_rate = ctcf_occ_rate) 

## Add E, P blocking
sim <- sim %>% add_ctcf(orientation = "+", pos_bp=c(819416), occupancy_rate = ctcf_occ_rate/2) 
sim <- sim %>% add_ctcf(orientation = "-", pos_bp=c(925000), occupancy_rate = ctcf_occ_rate/2) 

pointA <- 819 # 819416
pointB <- 925
# Full distance
pointB - pointA
sim_nt <- sim
sim_b <- sim_nt %>% delete_ctcf(pos_bp = c(910245, 915619, 919002))
sim_c <- sim_nt %>% delete_ctcf(pos_bp = c(888726, 910245, 915619, 919002))
pointA <- 819 # 819416 FGF3P
#pointA <- 887 # Left of boundary
pointB <- 925
if (file.exists(rds)){
  df <- readRDS(rds)
} else {
  num_iter <- 50000
  burn_in <- 10000
  num_reps <- 100
  
  system.time({
    df <- rbind(
      cbind(effective_dist_sim(sim, type="pairwise", pointA=pointA, pointB=pointB, num_iter=num_iter, num_reps=num_reps, burn_in=burn_in), cond="NT"),
      cbind(effective_dist_sim(sim_b, type="pairwise", pointA=pointA, pointB=pointB, num_iter=num_iter, num_reps=num_reps, burn_in=burn_in), cond="B"),
      cbind(effective_dist_sim(sim_c, type="pairwise", pointA=pointA, pointB=pointB, num_iter=num_iter, num_reps=num_reps, burn_in=burn_in), cond="C"))
  })
  
  df$cond <- factor(df$cond, levels=c("NT", "B", "C"))
  saveRDS(df, file=rds)  
}
```

## Plots

```{r}
df <- foreach (rds = dir("rds", pattern="*100reps.rds", full.names = TRUE), .combine=rbind) %do% {
  readRDS(rds) %>% cbind(sim=rds) 
}

df2 <- df %>% filter(effective_dist<=30)
df2 %>% filter(effective_dist<=20) %>% 
        mutate(num_cohesins = factor(str_split_fixed(sim, "-", 9)[,5], levels=c("coh8", "coh12", "coh16")),
               capture_rate = str_split_fixed(sim, "-", 9)[,7]) %>%
  group_by(num_cohesins, capture_rate, cond, effective_dist) %>% summarize(n=n()) %>% group_by(num_cohesins, capture_rate, cond) %>% mutate(cumsum = cumsum(n)) %>% ggplot(aes(effective_dist, cumsum, color=cond)) + geom_line() + xlim(0,15) + ylim(0, 30000) + theme_bw() +
  facet_grid(num_cohesins~capture_rate)
```



```{r}
df %>% filter(effective_dist<=20) %>% 
        mutate(num_cohesins = factor(str_split_fixed(sim, "-", 9)[,5], levels=c("coh8", "coh12", "coh16")),
               capture_rate = str_split_fixed(sim, "-", 9)[,7]) %>%
  group_by(num_cohesins, capture_rate, cond, effective_dist) %>% summarize(n=n()) %>% group_by(num_cohesins, capture_rate, cond)  %>% ggplot(aes(effective_dist, n, color=cond)) + geom_line() + xlim(0,5) + ylim(0, 3500) + theme_bw() +
  facet_grid(num_cohesins~capture_rate)

```

# Contact estimate
```{r}
k <- -1.5

df %>% mutate(num_cohesins = factor(str_split_fixed(sim, "-", 9)[,5], levels=c("coh8", "coh12", "coh16")),
               capture_rate = str_split_fixed(sim, "-", 9)[,7],
               num_contacts = effective_dist^k) %>%
  group_by(num_cohesins, capture_rate, cond) %>% summarize(n=n(), num_contacts=sum(num_contacts)) %>%
     ggplot(aes(cond, num_contacts, color=cond)) + geom_point()  + theme_bw() +
  facet_grid(num_cohesins~capture_rate)

```

```{r}
k <- -1.5

df %>% filter(effective_dist<=20) %>% 
    mutate(num_cohesins = factor(str_split_fixed(sim, "-", 9)[,5], levels=c("coh8", "coh12", "coh16")),
               capture_rate = str_split_fixed(sim, "-", 9)[,7],
               num_contacts = effective_dist^k) %>%
  group_by(num_cohesins, capture_rate, cond) %>% summarize(n=n(), num_contacts=sum(num_contacts)) %>%
     ggplot(aes(cond, num_contacts, color=cond)) + geom_point()  + theme_bw() +
  facet_grid(num_cohesins~capture_rate)
```


```{r}
df %>% ggplot(aes(effective_dist)) + geom_histogram(binwidth = 1) + facet_wrap(~cond, ncol=1) + theme_bw() + xlim(0, 105)
```

```{r}
df %>% group_by(cond, effective_dist) %>% summarize(n=n()) %>% group_by(cond) %>% mutate(cumsum = cumsum(n)) %>% ggplot(aes(effective_dist, cumsum, color=cond)) + geom_line() + xlim(0,30) + ylim(0, 60000) + theme_bw() 

```




The code below explores the required burn-in time and stability.

# Simulate

```{r}
system.time(
sim_df <- foreach(i = 1:100000, .combine=rbind) %do% {
        sim <- advance(sim)
        data.frame(i=i, effective_dist = effective_dist(sim, pointA, pointB), 
                   num_extruded_beads=sum(sim@extruded_beads),
                    bound_plus_cohesins=sum(sim@plus_ctcfs$bound_cohesin),
                    bound_minus_cohesins=sum(sim@minus_ctcfs$bound_cohesin))
})

sim_df %>% mutate(effective_dist_sim = cumsum(effective_dist)/i) %>% ggplot(aes(i, effective_dist_sim)) + geom_line() + theme_bw() + geom_hline(yintercept = pointB-pointA)
```

```{r}
sim_df %>% mutate(i_bin=round(i/5000)) %>% group_by(i_bin) %>% summarize(i = 5000*mean(i_bin), effective_dist = mean(effective_dist)) %>% ggplot(aes(i, effective_dist)) + geom_line() + theme_bw() + geom_hline(yintercept = pointB-pointA)

```


```{r}
sim_df %>% ggplot(aes(i, effective_dist)) + geom_line() + theme_bw()
```

```{r}
sim_df %>% ggplot(aes(effective_dist)) + geom_histogram() + theme_bw()

```


```{r}
sim_df %>% ggplot(aes(i, bound_minus_cohesins)) + geom_line() + theme_bw()
head(sim_df)

```

# Check simulation stability - No CTCFs
```{r}
sim <- Sim(8, polymer_length_bp = 240e3*8, processivity = 150e3, bead_size = 1000)

pointA <- 700
pointB <- 1200

sim_df <- foreach(run = 1:8, .combine=rbind) %dopar% {
    foreach(i = 1:50000, .combine=rbind) %do% {
        sim <- advance(sim)
        data.frame(run=run, i=i, effective_dist = effective_dist(sim, pointA, pointB))
    }
}
sim_df <- sim_df %>% group_by(run) %>% mutate(effective_dist_sim = cumsum(effective_dist)/i)
sim_df %>% ggplot(aes(i, effective_dist_sim, group=run)) + geom_line() + theme_bw()
#sim_df %>% filter(i>5000) %>% ggplot(aes(i, effective_dist_sim, group=run)) + geom_line() + theme_bw()
#saveRDS(sim_df, file="../data/sim_df.rds")
sim_df <- readRDS("../data/sim_df.rds")

```


```{r}

sim_df %>% filter(i>=150) %>% mutate(i=i-150) %>% group_by(run) %>% mutate(effective_dist_sim = cumsum(effective_dist)/i) %>% ggplot(aes(i, effective_dist_sim, group=run)) + geom_line() + theme_bw()

```


# Check simulation stability - With CTCFs
```{r}
sim <- sim_with_4_left_and_4_right_ctcf

pointA <- 700
pointB <- 1200

sim_df <- foreach(run = 1:4, .combine=rbind) %dopar% {
    foreach(i = 1:50000, .combine=rbind) %do% {
        sim <- advance(sim)
        data.frame(run=run, i=i, effective_dist = effective_dist(sim, pointA, pointB), 
                   num_extruded_beads=sum(sim@extruded_beads),
                    bound_plus_cohesins=sum(sim@plus_ctcfs$bound_cohesin),
                    bound_minus_cohesins=sum(sim@minus_ctcfs$bound_cohesin))
    }
}
saveRDS(sim_df, file="../data/sim_df_with_ctcf.rds")
sim_df <- readRDS("../data/sim_df_with_ctcf.rds")

```

# Effective length
```{r}
sim_df %>% filter(i>=150) %>% mutate(i=i-150) %>% group_by(run) %>% mutate(effective_dist_sim = cumsum(effective_dist)/i) %>% ggplot(aes(i, effective_dist_sim, group=run)) + geom_line() + theme_bw()

```
```{r}
sim_df %>% mutate(i_bin=5000*ceiling(i/5000)) %>% group_by(run, i_bin) %>% summarize(effective_dist_sim = mean(effective_dist))  %>% ggplot(aes(i_bin, effective_dist_sim, group=run)) + geom_line() + theme_bw() 

```


Why does it trend up?

# Num extruded beads
```{r}
sim_df %>% filter(i>=150) %>% mutate(i=i-150) %>% group_by(run) %>% mutate(avg_extruded_beads = cumsum(num_extruded_beads)/i) %>% ggplot(aes(i, avg_extruded_beads, group=run)) + geom_line() + theme_bw()

```




```{r}
sim_df %>% mutate(i_bin=5000*ceiling(i/5000)) %>% group_by(run, i_bin) %>% summarize(avg_extruded_beads = mean(num_extruded_beads))  %>% ggplot(aes(i_bin, avg_extruded_beads, group=run)) + geom_line() + theme_bw() 

```



# Num cohesin-bound CTCFs
```{r}
sim_df %>% mutate(i_bin=5000*ceiling(i/5000)) %>% group_by(run, i_bin) %>% summarize(bound_plus_cohesins = mean(bound_plus_cohesins),
                                                                                     bound_minus_cohesins = bound_minus_cohesins)  %>% 
    ggplot(aes(i_bin, bound_plus_cohesins, group=run)) + geom_line() + theme_bw() 

```
